---
title: "WebScrape"
author: "Nau, Gregory"
date: "April 11, 2019"
output: html_document
---

To run RSelenium I had to install an older version of Chromedriver and put it in my system directory.
I also had to uninstall all other Chromedriver versions, and Chrome in general for my computer (I shouldn't have needed to, but that is what got it to work)
So this code may or may not run on your machine.

Sets up the dataframe and library dependencies
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RSelenium)
library(xml2)
library(rvest)
library(dplyr)
library(naniar)

# Setting up columns
StructureOfFamilies <- data.frame(matrix(ncol = 10, nrow = 0))
colnames(StructureOfFamilies) <- c("Country", "Year", "Fertility", "AgeAtBirth", "OutOfWedlock", "Marriage", "Divorce", "TwoParents", "SingleParent", "OtherParent")
StructureOfFamilies <- StructureOfFamilies %>%
  mutate(Country = as.factor(Country)) %>%
  mutate(Year = as.factor(Year)) %>%
  mutate(Fertility = as.factor(Fertility)) %>%
  mutate(AgeAtBirth = as.factor(AgeAtBirth)) %>%
  mutate(OutOfWedlock = as.factor(OutOfWedlock)) %>%
  mutate(Marriage = as.factor(Marriage)) %>%
  mutate(Divorce = as.factor(Divorce)) %>%
  mutate(TwoParents = as.factor(TwoParents)) %>%
  mutate(SingleParent = as.factor(SingleParent)) %>%
  mutate(OtherParent = as.factor(OtherParent))
```

Does the webscraping
```{r webscrape}
# Opening Selinium remote server
rD <- rsDriver(chromever = "73.0.3683.68")
remDr <- rD$client
oecdURL <- "https://stats.oecd.org/index.aspx?queryid=68249"
remDr$navigate(oecdURL)

# loop through all countries in the OECD database
for (index in 1:50) {
  
  # index 37 = Non-OECD Economies, no data avaliable
  if (index %in% c(37)) {
    next()
  }
  
  # navigate to new country element
  countrySelector <- remDr$findElement(using = "id", "PDim_COU")
  countryOptions <- countrySelector$findChildElements(using = "tag name", "option")
  countryOptions[[index]]$clickElement()
  
  # clean environment
  rm(countrySelector)
  rm(countryOptions)
  
  # wait for table to load. User presses enter when loaded
  invisible(readline(prompt="Press [enter] to continue"))
  
  # do actual scraping
  
  # turns RSelenium page into HTML for RVest
  html <- read_html(remDr$getPageSource()[[1]])

  # get the avaliable years
  Year <- vector('character')
  y <- html %>%
    html_nodes(xpath = '//*[@class="DataTable"]//*[@class="HDim"]')
  num <- length(y)/2
  for (x in 1:num) {
    Year <- append(Year, html_text(y[[x]]))
  }
  rm(y)
  
  # pull this country name
  Country <- vector('character')
  c <- html %>%
    html_nodes(xpath = '//*[@id="PDim_COU"]/option')
  for (x in 1:num) {
    Country <- append(Country, html_text(c[[index]]))
  }
  rm(c)
  
  # get the avaliable fertility rates
  Fertility <- vector('character')
  f <- html %>%
    html_nodes(xpath = '//*[@id="row1"]//*[@class="Data"]')
  for (x in 1:num) {
    Fertility <- append(Fertility, html_text(f[[x]]))
  }
  rm(f)
  
  # get the age of mothers at birth
  AgeAtBirth <- vector('character')
  a <- html %>%
    html_nodes(xpath = '//*[@id="row2"]//*[@class="Data2"]')
  if (length(a) == 0) {
    for (x in 1:num) {
      AgeAtBirth <- append(AgeAtBirth, NA)
    }
  } else {
    for (x in 1:num) {
      AgeAtBirth <- append(AgeAtBirth, html_text(a[[x]]))
    }
  }
  rm(a)
  
  # get the unmarried births
  OutOfWedlock <- vector('character')
  o <- html %>%
    html_nodes(xpath = '//*[@id="row3"]//*[@class="Data"]')
  if (length(o) == 0) {
    for (x in 1:num) {
      OutOfWedlock <- append(OutOfWedlock, NA)
    }
  } else {
    for (x in 1:num) {
      OutOfWedlock <- append(OutOfWedlock, html_text(o[[x]]))
    }
  }
  rm(o)
  
  # get the marriage rate
  Marriage <- vector('character')
  m <- html %>%
    html_nodes(xpath = '//*[@id="row4"]//*[@class="Data2"]')
  if (length(m) == 0) {
    for (x in 1:num) {
      Marriage <- append(Marriage, NA)
    }
  } else {
    for (x in 1:num) {
      Marriage <- append(Marriage, html_text(m[[x]]))
    }
  }
  rm(m)
  
  # get the divorce rate
  Divorce <- vector('character')
  d <- html %>%
    html_nodes(xpath = '//*[@id="row5"]//*[@class="Data"]')
  if (length(d) == 0) {
    for (x in 1:num) {
      Divorce <- append(Divorce, NA)
    }
  } else {
    for (x in 1:num) {
      Divorce <- append(Divorce, html_text(d[[x]]))
    }
  }
  rm(d)
  
  # get the raised by 2 parents
  TwoParents <- vector('character')
  t <- html %>%
    html_nodes(xpath = '//*[@id="row6"]//*[@class="Data2"]')
  if (length(t) == 0) {
    for (x in 1:num) {
      TwoParents <- append(TwoParents, NA)
    }
  } else {
    for (x in 1:num) {
      TwoParents <- append(TwoParents, html_text(t[[x]]))
    }
  }
  rm(t)
  
  # get the raised by 1 parent
  SingleParent <- vector('character')
  s <- html %>%
    html_nodes(xpath = '//*[@id="row7"]//*[@class="Data"]')
  if (length(s) == 0) {
    for (x in 1:num) {
      SingleParent <- append(SingleParent, NA)
    }
  } else {
    for (x in 1:num) {
      SingleParent <- append(SingleParent, html_text(s[[x]]))
    }
  }
  rm(s)
  
  # get the raised by other parents
  OtherParent <- vector('character')
  n <- html %>%
    html_nodes(xpath = '//*[@id="row8"]//*[@class="Data2"]')
  if (length(n) == 0) {
    for (x in 1:num) {
      OtherParent <- append(OtherParent, NA)
    }
  } else {
    for (x in 1:num) {
      OtherParent <- append(OtherParent, html_text(n[[x]]))
    }
  }
  rm(n)
  
  # garbage collection
  rm(x)
  
  # add the data to the dataframe
  temp <- data.frame(Country, Year, Fertility, AgeAtBirth, OutOfWedlock, Marriage, Divorce, SingleParent, TwoParents, OtherParent)
  StructureOfFamilies <- full_join(StructureOfFamilies, temp)
  
  # garbage collection
  rm(temp)
  rm(AgeAtBirth)
  rm(Country)
  rm(Divorce)
  rm(Fertility)
  rm(Marriage)
  rm(OtherParent)
  rm(OutOfWedlock)
  rm(SingleParent)
  rm(TwoParents)
  rm(Year)
  rm(num)
  
}

# Closing Selinium remote server
rD[["server"]]$stop()

# garbage collection
rm(html)
rm(oecdURL)
rm(rD)
rm(remDr)
rm(index)
```

Cleans the oddities in the dataframe and saves it to a csv
```{r clean}
# fix China being weird (AgeAtBirth = Marriage, OutOfWedlock = Divorce)
china <- StructureOfFamilies %>%
  filter(grepl("China", Country))
StructureOfFamilies <- StructureOfFamilies %>%
  filter(!grepl("China", Country))
china <- china %>%
  rename("m" = AgeAtBirth) %>%
  rename("d" = OutOfWedlock) %>%
  rename("a" = Marriage) %>%
  rename("o" = Divorce)
china <- china %>%
  rename("Marriage" = m) %>%
  rename("Divorce" = d) %>%
  rename("AgeAtBirth" = a) %>%
  rename("OutOfWedlock" = o)
StructureOfFamilies <- full_join(StructureOfFamilies, china)
rm(china)

# turn "" into NA
StructureOfFamilies <- StructureOfFamilies %>%
  replace_with_na_all(condition = ~.x == "..")

# writes to csv for ease of stuff
write.csv(StructureOfFamilies, file = "StructureOfFamilies.csv")
```
